---
title: "referencemodel"
author: "zheng dong"
date: "2025-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = "hide", message = F, echo = F)
options(scientific=T, digits = 3) 
```

```{r init, include=FALSE}
# If you are missing any of the required packages, please install them first:
# install.packages(c("rmdformats","kableExtra","corrplot","GGally","caTools",
#                    "Metrics","glmnet","rpart.plot","tidymodels","caret",
#                    "randomForest","MASS","car","dplyr","ggplot2"))

knitr::opts_chunk$set(warning = FALSE, results = "hide", message = FALSE, echo = FALSE)
options(scientific = TRUE, digits = 3)
```

## 1. Loading Libraries and Reading the Dataset
```{r}
library(dplyr)
library(MASS)
library(car)
library(caTools)
library(Metrics)
library(glmnet)
library(ggplot2)
library(rpart)  # for decision trees
library(rpart.plot)
library(tidymodels)
library(tidyr)
library(caret)
library(knitr)
library(kableExtra)
library(corrplot)
library(GGally)
library(randomForest)

# Set the working directory (adjust as needed)
setwd("/Users/zhengdong/Documents/GitHub/Movie-Revenue-Prediction")

# Read in the CSV file. Make sure 'movies.csv' is in the working directory.
movies <- read.csv("final_dataset.csv")

movie  <- movies 

```


## 2. Preview of the Original Dataset

```{r}
# Display the first 5 rows to get a quick preview

head(movie, 5) %>%
  kable(caption = "The first 5 rows of the dataset") %>%
  kable_styling(full_width = FALSE)
```

##3. Removing Unnecessary Columns and Renaming
```{r}
remove_cols <- c("actor_2_name", "actor_1_name")  
movie <- movie[, !(names(movie) %in% remove_cols)]

movie <- movie %>% rename(
  Title = name,
  Genre = genre,
  Score = score,
  Director = director,
  Revenue = gross,
  Budget = budget
)

movie$Budget <- as.numeric(movie$Budget)
movie$Revenue <- as.numeric(movie$Revenue)
movie$Score <- as.numeric(movie$Score)


data_summary <- data.frame(
  Column = names(movie),
  Type = sapply(movie, class),
  Missing_Values = colSums(is.na(movie))
)

data_summary %>%
  kable(caption = "Dataset Summary (After Removing Some Columns)") %>%
  kable_styling(full_width = FALSE)


```

## 4. Handling Missing Values and Cleaning Invalid Rows

```{r}
column_na_percentages <- colSums(is.na(movie)) / nrow(movie) * 100

print("Rows in the dataset before dropping rows with NA values:")
print(nrow(movie))

movie <- movie[!is.na(movie$Revenue) & !is.na(movie$Budget), ]

print("Rows in the dataset after dropping rows with NA values:")
print(nrow(movie))

print("Rows in the dataset before cleaning:")
print(nrow(movie))

movie <- movie[movie$Budget > 0, ]
movie <- movie[movie$Revenue > 0, ]

print("Rows in the dataset after cleaning:")
print(nrow(movie))


```


## 5 Exploratory Data Analysis
## 5.1 Box Plots for Numeric Columns
```{r}
selected_columns <- c("Budget", "Revenue", "Score")  

selected_columns <- selected_columns[selected_columns %in% names(movie)]

box_plots_list <- list()
for (col in selected_columns) {
  box_plot <- ggplot(movie, aes(y = get(col))) +
    geom_boxplot(fill = "lightblue", color = "black", alpha = 0.7) +
    labs(title = paste("Box Plot for", col), y = col) +
    theme_minimal()
  box_plots_list[[col]] <- box_plot
}

for (col in selected_columns) {
  print(box_plots_list[[col]])
}


```

##5.2 Scatter Plots for Selected Predictors
```{r}
available_numeric_cols <- names(movie)[sapply(movie, is.numeric)]

predictors <- c("Budget", "Score") 
predictors <- predictors[predictors %in% available_numeric_cols]

scatter_plots <- lapply(predictors, function(var) {
  ggplot(movie, aes_string(x = var, y = "Revenue", color = "Revenue")) +
    geom_point() +
    labs(title = paste("Scatter plot of", var, "vs. Revenue"),
         x = var,
         y = "Revenue",
         color = "Revenue") +
    theme_minimal()
})
scatter_plots


```

##5.3 Correlation Matrix and Pairwise Plots
```{r}
numeric_cols <- sapply(movie, is.numeric)
numeric_vars_subset <- c("Revenue", "Budget", "Score")  

numeric_vars_subset <- numeric_vars_subset[numeric_vars_subset %in% names(movie)]

cor_matrix <- cor(movie[, numeric_cols])

corrplot(cor_matrix, method = "color")

if (length(numeric_vars_subset) > 1) {
  ggpairs(movie[, numeric_vars_subset])
} else {
  print("Not enough numeric variables for pairwise plotting.")
}

```

##6. Building the Initial Linear Models
##6.1 Stepwise Forward Selection
```{r}
library(MASS)


available_numeric_cols <- names(movie)[sapply(movie, is.numeric)]

potential_predictors <- c("Budget", "Score", "Revenue")  
potential_predictors <- potential_predictors[potential_predictors %in% available_numeric_cols]

if (length(potential_predictors) > 1) {
  formula_upper <- as.formula(paste("Revenue ~", paste(potential_predictors, collapse = " + ")))

  initial_model <- lm(Revenue ~ 1, data = movie)
  final_model <- stepAIC(initial_model, 
                         direction = "forward",
                         scope = list(lower = ~1, 
                                      upper = formula_upper),
                         data = movie)

  summary(final_model)
} else {
  print("Not enough numeric predictors for stepwise selection.")
}


```
##6.2 Fitting the Model Based on Stepwise Results
```{r}
selected_features <- c("Budget", "Score") 
selected_features <- selected_features[selected_features %in% names(movie)] 

if (length(selected_features) > 0) {
  formula_lm <- as.formula(paste("Revenue ~", paste(selected_features, collapse = " + ")))

  lm_model_usingFFS <- lm(formula_lm, data = movie)
  summary(lm_model_usingFFS)

  library(car)
  influencePlot(lm_model_usingFFS)
} else {
  print("Not enough predictors available for linear regression.")
}


```

##6.3 Removing Influential Points and Final Model
```{r}
rows_to_remove <- c(15996, 443, 96, 64, 1)
rows_to_remove <- rows_to_remove[rows_to_remove %in% rownames(movie)] 
movie <- movie[!(rownames(movie) %in% rows_to_remove), ]

if (length(selected_features) > 0) {
  lm_model <- lm(formula_lm, data = movie)
  summary(lm_model)
  influencePlot(lm_model)
} else {
  print("Not enough predictors available for final linear regression model.")
}


```


##7. Checking Assumptions of Linear Regression
```{r}
if (exists("lm_model") && !is.null(lm_model)) {
  
  vif_values <- vif(lm_model)
  print(vif_values)  
  
  avPlots(lm_model)
  
  par(mfrow = c(2, 2))  
  plot(lm_model)
  par(mfrow = c(1, 1))  

} else {
  print("Error: Linear model (lm_model) is not available.")
}


```

##8. Train-Test Split and Model Evaluation

```{r}
set.seed(123)
split <- sample.split(movie$Revenue, SplitRatio = 0.8)
train_data <- subset(movie, split == TRUE)
test_data  <- subset(movie, split == FALSE)

valid_columns <- c("Trailer_Views", "Trailer_Likes", "Vote_Count", "Budget")
valid_columns <- valid_columns[valid_columns %in% names(train_data)]  

if (length(valid_columns) > 0) {
    formula <- as.formula(paste("Revenue ~", paste(valid_columns, collapse = " + ")))
    lm_model <- lm(formula, data = train_data)
    
    train_predictions <- predict(lm_model, newdata = train_data)
    test_predictions  <- predict(lm_model, newdata = test_data)
    
    test_rmse  <- sqrt(mean((test_data$Revenue  - test_predictions)^2, na.rm = TRUE))
    train_rmse <- sqrt(mean((train_data$Revenue - train_predictions)^2, na.rm = TRUE))
    
    cat("Model Variables:", paste(valid_columns, collapse = ", "), "\n")
    cat("Test RMSE  :", round(test_rmse, 2), "\n")
    cat("Train RMSE :", round(train_rmse, 2), "\n")
    cat("Mean Revenue in dataset:", round(mean(movie$Revenue, na.rm=TRUE), 2), "\n")

} else {
    cat("Error: No valid predictor variables available for modeling.\n")
}


```
##9. Ridge and Lasso Regression
##9.1 Simple Ridge and Lasso with Lambda = 1
```{r}

valid_predictors <- c("Budget", "Score")  
valid_predictors <- valid_predictors[valid_predictors %in% names(train_data)]

if (length(valid_predictors) >= 2) {
  X_train <- as.matrix(train_data[valid_predictors])
  X_test  <- as.matrix(test_data[valid_predictors])
  y_train <- train_data$Revenue
  y_test  <- test_data$Revenue

  ridge_model <- glmnet(X_train, y_train, alpha = 0, lambda = 1)
  lasso_model <- glmnet(X_train, y_train, alpha = 1, lambda = 1)

  ridge_train_predictions <- predict(ridge_model, newx = X_train)
  ridge_test_predictions  <- predict(ridge_model, newx = X_test)
  lasso_train_predictions <- predict(lasso_model, newx = X_train)
  lasso_test_predictions  <- predict(lasso_model, newx = X_test)

  ridge_train_rmse <- sqrt(mean((y_train - ridge_train_predictions)^2, na.rm = TRUE))
  ridge_test_rmse  <- sqrt(mean((y_test - ridge_test_predictions)^2, na.rm = TRUE))
  lasso_train_rmse <- sqrt(mean((y_train - lasso_train_predictions)^2, na.rm = TRUE))
  lasso_test_rmse  <- sqrt(mean((y_test - lasso_test_predictions)^2, na.rm = TRUE))

  cat("Final Predictors Used:", paste(valid_predictors, collapse = ", "), "\n")
  cat("Ridge Train RMSE:", round(ridge_train_rmse, 2), "\n")
  cat("Ridge Test RMSE:", round(ridge_test_rmse, 2), "\n")
  cat("Lasso Train RMSE:", round(lasso_train_rmse, 2), "\n")
  cat("Lasso Test RMSE:", round(lasso_test_rmse, 2), "\n")
} else {
  cat("Error: Not enough valid predictor variables for Ridge/Lasso Regression.\n")
}

```
##9.2 Cross-Validated Ridge and Lasso

```{r}
set.seed(42)

valid_predictors <- c("Budget", "Score")  
valid_predictors <- valid_predictors[valid_predictors %in% names(train_data)]

if (length(valid_predictors) >= 2) {
  X_train <- as.matrix(train_data[valid_predictors])
  X_test  <- as.matrix(test_data[valid_predictors])
  y_train <- train_data$Revenue
  y_test  <- test_data$Revenue

  lambdas <- 10^seq(10, -1, length = 1000)

  ridge_cv_model <- cv.glmnet(X_train, y_train, alpha = 0, lambda = lambdas)
  lasso_cv_model <- cv.glmnet(X_train, y_train, alpha = 1, lambda = lambdas)

  best_lambda_ridge <- ridge_cv_model$lambda.min
  best_lambda_lasso <- lasso_cv_model$lambda.min

  cat("Best lambda (ridge):", best_lambda_ridge, "\n")
  cat("Best lambda (lasso):", best_lambda_lasso, "\n")

  ridgetuned_train_predictions <- predict(ridge_cv_model, s = best_lambda_ridge, newx = X_train)
  ridgetuned_test_predictions  <- predict(ridge_cv_model, s = best_lambda_ridge, newx = X_test)
  lassotuned_train_predictions <- predict(lasso_cv_model, s = best_lambda_lasso, newx = X_train)
  lassotuned_test_predictions  <- predict(lasso_cv_model, s = best_lambda_lasso, newx = X_test)

  ridgetuned_train_rmse <- sqrt(mean((y_train - ridgetuned_train_predictions)^2, na.rm = TRUE))
  ridgetuned_test_rmse  <- sqrt(mean((y_test - ridgetuned_test_predictions)^2, na.rm = TRUE))
  lassotuned_train_rmse <- sqrt(mean((y_train - lassotuned_train_predictions)^2, na.rm = TRUE))
  lassotuned_test_rmse  <- sqrt(mean((y_test - lassotuned_test_predictions)^2, na.rm = TRUE))

  cat("Final Predictors Used:", paste(valid_predictors, collapse = ", "), "\n")
  cat("Ridge (tuned) Train RMSE:", round(ridgetuned_train_rmse, 2), "\n")
  cat("Ridge (tuned) Test  RMSE:", round(ridgetuned_test_rmse, 2),  "\n")
  cat("Lasso (tuned) Train RMSE:", round(lassotuned_train_rmse, 2), "\n")
  cat("Lasso (tuned) Test  RMSE:", round(lassotuned_test_rmse, 2),  "\n")

} else {
  cat("Error: Not enough valid predictor variables for Ridge/Lasso Regression.\n")
}

```

##10. Decision Trees and Random Forest
##10.1 Decision Tree (Model 0: All Features)
```{r}
set.seed(1)

valid_features <- c("Budget", "Score", "Vote_Count")
valid_features <- valid_features[valid_features %in% names(train_data)]

train_data <- na.omit(train_data)
test_data <- na.omit(test_data)
if (length(valid_features) > 1) {
  dt_model0 <- rpart(as.formula(paste("Revenue ~", paste(valid_features, collapse = " + "))),
                     data = train_data, method = "anova")
  
  y_pred <- predict(dt_model0, test_data)
  cat("DT Model0 RMSE:", RMSE(test_data$Revenue, y_pred), "\n")
  printcp(dt_model0)
  rpart.plot(dt_model0)
  plotcp(dt_model0)
} else {
  cat("Error: Not enough valid predictors for Decision Tree Model 0.\n")
}

```


